---
title: '2022 Midterm Forecast Introduction'
author: Corine Chung
date: '2022-09-10'
slug: []
categories: []
tags: []
summary: "From now until November 8, I will be updating this weekly blog series with my **2022 US midterm election prediction model**. For this first blog, I'll be exploring past election results to find any trends in the data. More specifically, I'll be diving into the state of **Michigan**."
---

```{r setup, include=FALSE}
# Hiding all code output
knitr::opts_chunk$set(echo = FALSE)
# Loading libraries
library(tidyverse)
library(sf)
library(plotly)
# load usmap
# install.packages('plot_usmap')
library(usmap)
h <- read_csv("house party vote share by district 1948-2020.csv")

```

*This blog is part of a series related to Gov 1347: Election Analytics, a course at [Harvard University](https://www.harvard.edu/) taught by Professor [Ryan D. Enos](http://ryandenos.com/)*.

<br>

From now until November 8, I will be updating this weekly blog series with my **2022 US midterm election prediction model**. For this first blog, I'll be exploring past election results to find any trends in the data. More specifically, I'll be diving into the state of **Michigan**.

<br>

# United States

<br>

Before focusing specifically on Michigan, I first examine the GOP two-party vote share margin for the 2010 midterm election.

```{r}
## make map of vote share by state and CD

# start with 114th congress - 2014 election
# required packages 
require(tidyverse)
require(ggplot2)
require(sf)

# load 114th congress
cd114 <- st_read("districtShapes/districts114.shp", quiet = T)

# select specific state 
cd114_mi <- cd114 %>% 
            filter(STATENAME=="Michigan") %>%
            mutate(DISTRICT = as.character(DISTRICT))%>%
            select(DISTRICT)

# add data to plot - 2014 GOP party seat share
# reload election data - h from previous exercise
```


```{r}
## make map of vote share by state and CD

# start with 114th congress - 2014 election
# required packages 
require(tidyverse)
require(ggplot2)
require(sf)

# load 114th congress
cd114 <- st_read("districtShapes/districts114.shp", quiet = T)

# select specific state 
cd114_mi <- cd114 %>% 
            filter(STATENAME=="Michigan") %>%
            mutate(DISTRICT = as.character(DISTRICT))%>%
            select(DISTRICT)

## make map of GOP vote share by state (national) - 2014
# use h dataset from earlier
# house_party_vote_share_by_district_1948_2020 <- 
  # read_csv("house party vote share by district 1948-2020.csv")
# h <- house_party_vote_share_by_district_1948_2020
 
# filter for relevant variables
R_2010 <- h %>%
    filter(raceYear == 2010) %>%
    select(raceYear, State, district_num, district_id, RepVotes, DemVotes) %>%
  # summarize party vote share by state
    group_by(State) %>%
  # mutate Rep vote margin by state %>%
    mutate(R_votemargin_st = (sum(RepVotes))/
                            sum(RepVotes + DemVotes),
           D_votemargin_st = (sum(DemVotes))/
                            sum(RepVotes + DemVotes)) %>%
  rename(state = State)

states_map <- usmap::us_map()
unique(states_map$abbr)

# plot
plot_usmap(data = R_2010, regions = "states", values = "R_votemargin_st", labels = TRUE) + 
  scale_fill_gradient(low = "white", high = "red", name = "GOP two-party voteshare margin") +
  theme_void()
```
As determined by the key on the right, Michigan seemed to be a highly-contested state in 2010.

<br>

Next, I examine Michigan specifically, which has been a critical swing state in recent U.S. [presidential elections](https://www.270towin.com/states/Michigan). 

<br>

```{r}
R_mi_2014 <- h %>%
    filter(raceYear == 2014, State == "Michigan") %>%
    select(raceYear, State, district_num, RepVotesMajorPercent, DemVotesMajorPercent) %>%
  # summarize party vote share by district
    group_by(district_num) %>%
    summarise(Rep_votes_pct = RepVotesMajorPercent) %>%
  # rename district variable name to match shapefile
    rename(DISTRICT = district_num)

# before joining dfs, check classes of variable to be merged on
class(R_mi_2014$DISTRICT)
class(cd114_mi$DISTRICT)

# change class
cd114_mi$DISTRICT <- as.numeric(cd114_mi$DISTRICT)

# join election returns with shapefiles
cd114_mi <- cd114_mi %>% left_join(R_mi_2014, by="DISTRICT")
cd114_mi

ggplot() +
geom_sf(data=cd114_mi,aes(fill=Rep_votes_pct),
inherit.aes=FALSE,alpha=0.9) +
scale_fill_gradient(low = "white", high = "black", limits=c(10,80)) +
theme_void() +
theme(axis.title.x=element_blank(),
axis.text.x=element_blank(),
axis.ticks.x=element_blank(),
axis.title.y=element_blank(),
axis.text.y=element_blank(),
axis.ticks.y=element_blank())
``` 

